Grounding the interaction: anchoring situated discourse in everyday human-robot interaction

### 2011-07-15 paper_knowledge.tex
Integrated Raquel changes on SPARK

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic assessments. In this section we present SPARK (\emph{SPAtial
Reasoning \& Knowledge}), a situation assessment reasoner that generates
relevant symbolic information from the geometry of the environment with respect
to relations between objects, robots and humans. Moreover, the notion of
\emph{Perspective Taking}~\cite{Flavell1992,Tversky1999} is employed in the
heart of the reasoner to provide the robot with the ability to put itself at
the human’s place and to reason about the world from different perspectives.

@@

There are a number of common properties for a robot and a human related to
their capabilities in a given situation: they can both reach, grasp, look,
point, etc. In our context, we group the robot and the human into a single
category. Thus, we define the ``agent'' as an entity that can act in the
environment and manipulate it. In this work we focus on the following
capabilities from each agents' perspective \footnote{Note that each of the
capabilities described are computed from each agent point of view, and
therefore, also stored in different models in ORO for further use at the
decisional level.}.

@@

\item \emph{Looks At}: this relation corresponds to what the agent is focused
on, \ie, where its focus of attention is directed to. This model is based on a
narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
shows the field of attention of a person with a green cone (narrower one). In
this example only the grey box satisfies the \verb"looksAt" relation.

@@

If a big object occludes a smaller one, and an agent is pointing at them, the
outcome of the evaluation will result only in one relation, \ie, \stmt{Agent
\textbf{pointsAt} BigObject} since the small one is not visible to the agent.
On the contrary, if the small object is in front of the big one, then both
objects will be pointed at by the human, generating an ambiguity which should
be solved through higher level reasoning (\eg context analysis or
clarification through verbal interaction as shown later on).

@@

While the first three relations (\emph{sees, looksAt} and \emph{pointsAt}) are
computed through a model based approach, the latter one is based on the
Generalized Inverse Kinematics with pseudo inverse
method~\cite{Nakamura90,Baerlocher04} to find a collision free posture for the
agent where its end-effector is at the center of the object within a given
tolerance. The details of these computations are out of the scope of this
article.

\subsection{Locations} As humans, one way of referring to the object's
positions is based on their symbolic descriptors instead of using their precise
position. In fact, in many cases, this information is the most precise one that
we can give since we do not store the numeric coordinates of objects. The
following relations are computed with respect to the position of the agents and
the objects.

\begin{itemize} \item \emph{Location according to an agent} The predicate
\verb"isLocatedAt" represents spatial locations between agents and objects. For
example we say ``it is on my right, on your left,...'' We compute these spatial
locations by dividing the space around the referent (an agent) into $n$ regions
based on arbitrary angle values relative to the referent orientation. For
example, for $n = 4$ we would have the space divided into \emph{front, left,
right} and \emph{back}. Additionally, two proximity values, \emph{near} and
\emph{far}, may also be considered. The number of regions and proximity values
can be chosen depending on the context where the interaction takes place.

\item \emph{Location according to an object} We can also refer to object
locations with respect to other objects in the environment, such as, above,
next to, in, etc. These types of relations are widely studied in language
grounding (e.g.~\cite{Regier01} presented different models to define the
\emph{above} relation). In this work we use similar models based on the
bounding boxes and center of mass of the objects to define three main relations
(Figure~\ref{fig::sprelation}): \begin{itemize}

\item \textit{isOn}: computes if an object $O_1$ is on another object $O_2$ by
evaluating the center of mass of $O_1$ according to the bounding box of $O_2$.

\item \textit{isIn}: evaluates if an object $O_1$ is inside another object
$O_2$ based on their bounding boxes $BB_{O_1}$ and $BB_{O_2}$.

\item \textit{isNextTo}: indicates whether an object $O_1$ is next to another
object $O_2$. We cannot use a simple distance threshold to determine if two
objects are next to each other since the relation is highly dependent on the
dimensions of the objects. For instance, the maximum distance between large
objects (e.g. two houses) to consider them as being next to each other is much
larger than the maximum distance we would consider for two small objects (\eg
two bottles). Thus, the relation between the dimensions and the distances of
the objects are taken into account.

@@

SPARK can be compared to the \emph{Grounded Situation Model} (GSM) by
Mavridis and Roy~\cite{Mavridis2005} in the sense that they both provide an amodal physical
representation of the world used as mediator between the sensor space and
symbolic models. They have however a different set of features: while GSM
enable representation of time and imaginary objects (whose existence is hinted
by verbal assertions from a human), SPARK offers a richer 3D model that enable
computation of several spatial relationships between objects and an effective
implementation of the perspective taking.

>>>>

Anchoring perceptions in a symbolic model requires first perception abilities
and second, a symbolic assessment of the percepts. In this section we present
SPARK (\emph{SPAtial Reasoning \& Knowledge}), a situation assessment reasoner
that generates relevant symbolic information from the geometry of the
environment with respect to relations between objects and human capabilities.
Moreover, the notion of \emph{Perspective
Taking}~\cite{Flavell1992,Tversky1999} is employed in the heart of the
reasoning to provide the robot with the ability to put itself at the human's
place and to reason about the world from different perspectives.

@@

There are a number of common properties for a robot and a human related to
their capabilities in a given situation: they can both reach, grasp, look,
point, etc. In our context, we group robots and humans into a single category.
Thus, we define agents as entities that can act in the environment and
manipulate it. In this work we focus on the following capabilities from each
agent's perspective\footnote{Note that each of the capabilities described are
computed from each agent point of view, and therefore, also stored in different
models in ORO for further use at the decisional level.}.

@@

\item \emph{Looks At}: this relation corresponds to what the agent is focused
on, \ie where its focus of attention is directed to. This model is based on a
narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
shows the field of attention of a person with a green cone (narrower one). In
this example only the grey box satisfies the \concept{looksAt} relation.

@@

If a big object occludes a smaller one, and an agent is pointing at them, the
outcome of the evaluation will result only in one relation, \ie \stmt{agent\_01
pointsAt object\_01} since the small one is not visible to the agent.  On the
contrary, if the small object is in front of the big one, then both objects
will satisfy the relation, which may generate an ambiguity (which object the
agent refers to?) that should be solved through higher level reasoning (\eg
context analysis or clarification through verbal interaction as shown later
on).

@@

While the first three relations (\concept{sees}, \concept{looksAt} and
\concept{pointsAt}) are computed through a model based approach, the latter one
is based on the Generalized Inverse Kinematics with pseudo inverse
method~\cite{Nakamura90,Baerlocher04} to find a collision free posture for the
agent where its end-effector is at the center of the object within a given
tolerance. The details of these computations are out of the scope of this
article.

\subsection{Locations} As humans, one way of referring to object's
positions is based on their symbolic descriptors instead of using their precise
position. In fact, in many cases, this information is the most precise one that
we can give since we do not store the numeric coordinates of objects. The
following relations are computed with respect to the position of the agents and
the objects.

\item \emph{Location according to an agent}: The predicate
\concept{isLocatedAt} represents spatial locations between agents and objects.
For example we say ``it is on my right, on your left,...'' We compute these
spatial locations by dividing the space around the referent (an agent) into $n$
regions based on arbitrary angle values relative to the referent orientation.
For example, for $n = 4$ we would have the space divided into \emph{front,
left, right} and \emph{back}. Additionally, two proximity values, \emph{near}
and \emph{far}, may also be considered. The number of regions and proximity
values can be chosen depending on the context where the interaction takes
place.

\item \emph{Location according to an object}: We can also refer to object
locations with respect to other objects in the environment, such as \emph{above,
next to, in}, etc. These types of relations are widely studied in language
grounding (\eg \cite{Regier01} presented different models to define the
\emph{above} relation). In this work we use similar models based on the
bounding boxes and center of mass of the objects to define three main relations
(Figure~\ref{fig::sprelation}):

\begin{itemize}
	\item \concept{isOn}: computes if an object $O_1$ is on another object $O_2$ by
	evaluating the center of mass of $O_1$ according to the bounding box of $O_2$.

	\item \concept{isIn}: evaluates if an object $O_1$ is inside another object
	$O_2$ based on their bounding boxes $BB_{O_1}$ and $BB_{O_2}$.

	\item \concept{isNextTo}: indicates whether an object $O_1$ is next to another
	object $O_2$. We cannot use a simple distance threshold to determine if two
	objects are next to each other since the relation is highly dependent on the
	dimensions of the objects. For instance, the maximum distance between large
	objects (\eg two houses) to consider them as being next to each other is much
	larger than the maximum distance we would consider for two small objects (\eg
	two bottles). Thus, the relation between the dimensions and the distances of
	the objects are taken into account.

@@

SPARK can be compared to the \emph{Grounded Situation Model} (GSM) introduced
by Mavridis and Roy~\cite{Mavridis2005} in the sense that they both provide an
amodal physical representation of the world used as a mediator between the
sensor space and symbolic models. They have however a different set of
features: while GSM enables representation of time and imaginary objects (whose
existence is hinted by verbal assertions from a human), SPARK offers a richer
3D model that enables the computation of several spatial relationships between
objects and an effective implementation of perspective taking capabilities.
